---
title: "1-T-test"
author: "Mayla Boguslav"
date: "12/6/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## T-test background information 
Source: [Wikipedia](https://en.wikipedia.org/wiki/Student%27s_t-test)
As an example, in the one-sample t-test $t=\frac{Z}{s} = \frac{\frac{(\bar X-\mu)}{\frac{(\sigma)}{\sqrt n}}}{s}$
where $\bar X$ is the sample mean from a sample $X_1,X_2,…,X_n$, of size n, s is the ratio of sample standard deviation over population standard deviation, σ is the population standard deviation of the data, and μ is the population mean.


The assumptions underlying a t-test are:

* X follows a normal distribution with mean μ and variance σ^2^
* ps^2^ follows a χ^2^ distribution with p degrees of freedom under the null hypothesis, where p is a positive constant
* Z and s are independent.



## Import Data iris

```{r data_import}
data("iris")
data.subset<-subset(iris, iris[,5] %in% c("setosa", "versicolor")) #subset the data into only 2 species
# T-test to find differences in sepal length between setosa and versicolor species
observed.t = t.test(data.subset$Sepal.Length,data.subset$Species=="setosa")$statistic

```

## Permutation test
What if there is some correlation structure that is influencing our p-value?! What do we do?? 
Answer: a permutation test!!!! (see overview for details)

```{r perm-test}
#Function to permute class labels and outputs the t_value
perm.t = function(doesnt.matter, x, y){
  # Shuffle the species label
  y.shuffle = sample(y)
  # Calculate t-test on this new permutation
  t.value = t.test(x,y.shuffle)$statistic
  #print(y.shuffle==y) #check that the shuffling is different than the original
  return(t.value)
}

n = 10 #number of permutations
all.the.ts = sapply(1:n, perm.t, x=data.subset$Sepal.Length, y=data.subset$Species=='setosa')

# Check to see where our observed t_value sits on the distribution of permuted values
p.value = sum(all.the.ts <= abs(observed.t))/n #the p-value is significant if less than 0.05 generally


```




